POPLMark goes Concurrent!\\Challenge problems for process calculi and behavioural types

### 2023-10-12 challenges/linearity.tex
putting quad in bnf as in coinduction.tex
Fix linearity challenge introduction.

\begin{metanote}
 AM: do we mean type safety? If so the description is half the story
\end{metanote}
The objective of this challenge is to prove type preservation (also
known as subject reduction) for the type system, i.e.\ that well-typed
processes can only transition to processes which are also well-typed
in the same context.

\subsubsection{Syntax.}
The syntax is given by the grammar
\[
\begin{array}{rcl}
  v,w & ::= & a \mid l \\
   P,Q & ::=& \Pend \mid \Pout{x}{v}{P} \mid \Pin{x}{l}{P} \mid \PBpar{P}{Q} \mid  \Presd{x}{y}{P}
\end{array}
\]
where a \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).

@@

\paragraph{Syntax}
Our type system does not type processes directly, but instead focuses on the channels used in the process.
The syntax of \emph{session types} \( S, T \), unrestricted type contexts \( \Gamma \) and linear type contexts \( \Delta \) is as follows:
\[
  \begin{array}{rcl}
  S,T & ::= & \Tend \mid \Tbase \mid \Tin{S} \mid \Tout{S} \\
    \Gamma &::= & \Cempty \mid \Gamma, l \\
                    \Delta &::= & \Cempty \mid \Cadd{\Delta}{\hastype{x}{S}}
  \end{array}
\]
The \emph{end type} \( \Tend \) describes an endpoint on which no further interaction is possible.
The \emph{base type} \( \Tbase \) describes base values.
The \emph{input type} \( \Tin{S} \) describes endpoints used for receiving a value and then according to \( S \).
The \emph{output type} \( \Tout{S} \) describes endpoints used for sending a value and then according to \( S \).

>>>>

The main objective of this challenge is to prove type preservation (also
known as subject reduction), i.e.\ that well-typed
processes can only transition to processes which are also well-typed
in the same context.
The second objective is to prove type safety, i.e.\ that well-typed processes are also well-formed in the sense that they do not use endpoints in a non-dual way.

\subsubsection{Syntax.}
The syntax is given by the grammar
\[
\begin{array}{rcl}
  v,w & ::= & a \quad\mid\quad l \\
   P,Q & ::=& \Pend \quad\mid\quad \Pout{x}{v}{P} \quad\mid\quad \Pin{x}{l}{P} \quad\mid\quad \PBpar{P}{Q} \quad\mid\quad  \Presd{x}{y}{P}
\end{array}
\]
where a \emph{value} \( v, w, \dots \) is either a base value \( a \) or a variable name \( l \).

@@

\paragraph{Syntax}
Our type system does not type processes directly, but instead focuses on the channels used in the process.
The syntax of \emph{session types} \( S, T \), unrestricted type contexts \( \Gamma \) and linear type contexts \( \Delta \) is as follows:
\[
  \begin{array}{rcl}
  S,T & ::= & \Tend \quad\mid\quad \Tbase \quad\mid\quad \Tin{S} \quad\mid\quad \Tout{S} \\
    \Gamma &::= & \Cempty \quad\mid\quad \Gamma, l \\
                    \Delta &::= & \Cempty \quad\mid\quad \Cadd{\Delta}{\hastype{x}{S}}
  \end{array}
\]
The \emph{end type} \( \Tend \) describes an endpoint on which no further interaction is possible.
The \emph{base type} \( \Tbase \) describes base values.
The \emph{input type} \( \Tin{S} \) describes endpoints used for receiving a value and then according to \( S \).
The \emph{output type} \( \Tout{S} \) describes endpoints used for sending a value and then according to \( S \).
